---
title: "STAT/MATH 495: Problem Set 05"
author: "Jenn, Pei, and Anthony"
date: "2017-10-11"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    collapsed: false
    smooth_scroll: false
    df_print: kable
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, fig.width=8, fig.height=4.5, message=FALSE, warning = FALSE
  )
set.seed(76)

# Load packages
library(tidyverse)
library(corrplot)
library(broom)

# Note the relative file path, and not absolute file path:
train <- read_csv("data/train.csv")
test <- read_csv("data/test.csv")
sample_submission <- read_csv("data/sampleSubmission.csv")
```


# Collaboration

Please indicate who you collaborated with on this assignment:

For lm model fitting using dplyr: https://cran.r-project.org/web/packages/broom/vignettes/broom_and_dplyr.html

# Exploratory Data Analysis

Information on the competition can be found [here](https://www.kaggle.com/c/bike-sharing-demand).

```{r}
#Storing a version of time in system time 
train <- train %>% mutate(SysTime = as.numeric(unclass(datetime)))
test <- test %>% mutate(SysTime = as.numeric(unclass(datetime)))
#Making a corrlation plot of the variables, excluding registered and casual
#since count is the sum of registered and casual

#Note: Since count is our outcome variable, we are interested in variables that have
#a high correlation with count
corPlotData <- train %>% select(season, holiday, workingday, weather, temp, atemp, humidity, windspeed,count, SysTime) %>% cor()
corrplot(corPlotData, order = "hclust")

```

From the correlation plot, temperature, dateTime (SysTime in the plot), and humidity have the strongest relationship with bike rentals.

# Model Fit

Fitting the model to the training data, we can that each chosen predictor has a significant p-value.
```{r}
model <- train %>% do(fit = lm(count~ SysTime + temp + humidity, .))
model %>% tidy(fit)
```


# Create Submission File

Here, we've added a second layer to our model. If the model predicts bike rental numbers less than 0, we've set the estimate to 0.
```{r}
fit <- train %>% do(augment(lm(count~SysTime + temp + humidity,data= .),newdata = test))
fit$.fitted[fit$.fitted < 0] <- 0
dfToSubmit <- fit %>% mutate(count = .fitted) %>% select(datetime, count)

write.csv(dfToSubmit, file = "submission.csv", row.names = FALSE)
```
